---
title: "Wrangling"
author: "Candelaria Retamal"
date: "2023-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(devtools)
```

```{r}
drinks_data <- read_csv(here("data-raw/all_drinks.csv"))
```

We will start by checking the unique values contained in the columns that describe the ingredients

```{r}
# Get unique values in all ingredient columns
unique_ingredients <- unique(c(
  drinks_data$strIngredient1, drinks_data$strIngredient2, drinks_data$strIngredient3,
drinks_data$strIngredient4,
drinks_data$strIngredient5,
drinks_data$strIngredient6,
drinks_data$strIngredient7,
drinks_data$strIngredient8,
drinks_data$strIngredient9,
drinks_data$strIngredient10,
drinks_data$strIngredient11,
drinks_data$strIngredient12,
drinks_data$strIngredient13,
drinks_data$strIngredient14,
drinks_data$strIngredient15
))

# Define a function to replace specified ingredients
replace_ingredients <- function(x) {
  x <- tolower(x)
  x <- trimws(x)
  
  # Replace specified ingredients
  x <- ifelse(x %in% c("anis", "anise"), "anise", x)
  x <- ifelse(x %in% c("bailey's irish cream", "baileys irish cream"), "baileys irish cream", x)
  x <- ifelse(x %in% c("whipped cream", "whipped Cream", "whipping cream"), "whipped cream", x)
  x <- ifelse(x %in% c("whiskey", "whisky"), "whiskey", x)
  
  return(x)
}

# Apply the replacement function to all columns
drinks_data <- mutate_all(drinks_data, replace_ingredients)

# Convert to lowercase for standardization
unique_ingredients <- tolower(unique_ingredients)

# Remove leading and trailing whitespaces
unique_ingredients <- trimws(unique_ingredients)

# Get unique standardized values
unique_ingredients <- unique(unique_ingredients)

# Order the unique ingredient names alphabetically
unique_ingredients <- sort(unique_ingredients)

unique_ingredients <- sort(unique_ingredients)
#unique_ingredients
```

In the different steps we have checked the list of unique values. First, we solved the problems of words that differed in capitalization and number of spaces. Then we solved by hand words that belonged to a same root, but were written differently in the dataset. We are now sure that we have reduced the number of ingredients to the minimum possible.

```{r}
# Create a copy of the original dataset
drinks_data_combined <- drinks_data

# Combine corresponding Ingredient and Measure columns
for (i in 1:15) {
  ingredient_col <- paste0("strIngredient", i)
  measure_col <- paste0("strMeasure", i)
  
  new_col_name <- paste0("Ingredient_Measure_", i)
  
  drinks_data_combined[[new_col_name]] <- paste(drinks_data_combined[[ingredient_col]], drinks_data_combined[[measure_col]], sep = " ; ")
}

# Optionally, you can remove the individual ingredient and measurement columns
drinks_data_combined <- select(drinks_data_combined, -starts_with("strIngredient"), -starts_with("strMeasure"))

# Verify the changes
head(drinks_data_combined)
```

```{r}
# Create a copy of the original dataset
drinks_data_encoded <- drinks_data

# Get unique standardized values of ingredients
unique_ingredients <- c(
  drinks_data$strIngredient1, drinks_data$strIngredient2, drinks_data$strIngredient3,
  drinks_data$strIngredient4, drinks_data$strIngredient5, drinks_data$strIngredient6,
  drinks_data$strIngredient7, drinks_data$strIngredient8, drinks_data$strIngredient9,
  drinks_data$strIngredient10, drinks_data$strIngredient11, drinks_data$strIngredient12,
  drinks_data$strIngredient13, drinks_data$strIngredient14, drinks_data$strIngredient15
) %>%
  tolower() %>%
  trimws() %>%
  unique()

# Create one-hot encoded columns for each ingredient
for (ingredient in unique_ingredients) {
  column_name <- paste0("ingredient_", gsub(" ", "_", ingredient))
  drinks_data_encoded[column_name] <- as.integer(
    rowSums(sapply(1:15, function(i) grepl(ingredient, drinks_data[[paste0("strIngredient", i)]], ignore.case = TRUE))) > 0
  )
}

# Optionally, you can remove the individual ingredient columns
drinks_data_encoded <- select(drinks_data_encoded, -starts_with("strIngredient"))

# Verify the changes
head(drinks_data_encoded)

```

Create one column per ingredient and put the measurements instead of the binary 1 and 0. If it is a 1, replace for the actual measurement of the ingredient per drinking.

```{r}
for (row_index in 1:nrow(drinks_data_combined)) {
  # Iterate over columns Ingredients_Measures_X
  for (i in 1:15) {
    col_name <- paste("Ingredient_Measure_", i, sep = "")
    
    #Obtain entry of the current column
    entry <- as.character(drinks_data_combined[row_index, col_name])
    
    # Verify if entry is NA and of type chr
    if (!is.na(entry) && length(entry) > 0) {
      # Separate ingredient and measurement
      parts <- strsplit(entry, "; ", fixed = TRUE)[[1]]
      
      # Obtain ingredient and measurement
      ingredient <- parts[1]
      measure <- parts[2]
      
      # Value of the column that corresponds to ingredient and current row
      drinks_data_combined[row_index, ingredient] <- measure
    }
  }
}

# Verify result
head(drinks_data_combined)

```

Remove columns that have been wrangled using the binary approach. Also, remove those columns that do not contain enough or valuable information.

```{r}
# Eliminar columnas que contienen "Ingredient_Measure" en el nombre
columns_to_remove <- grep("Ingredient_Measure", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]

columns_to_remove <- grep("...1", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]

columns_to_remove <- grep("strIBA", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]

columns_to_remove <- grep("dateModified", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]

columns_to_remove <- grep("idDrink", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]

columns_to_remove <- grep("strVideo", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]

columns_to_remove <- grep("NA", names(drinks_data_combined))
drinks_data_combined <- drinks_data_combined[, -columns_to_remove]
head(drinks_data_combined)
```

Change the names of the columns, and rearrange the information correctly to make sure that the cleaning process is complete.

```{r}
# Rename the column
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "strDrink"] <- "Name"
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "strAlcoholic"] <- "Type"
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "strCategory"] <- "Category"
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "strDrinkThumb"] <- "Picture"
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "strGlass"] <- "Glass"
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "strInstructions"] <- "Recipe"
colnames(drinks_data_combined)[colnames(drinks_data_combined) == "151 proof rum "] <- "151 bacardi rum"
```

Eliminate the inconsistencies found in the dataset, on the section of the measurements.

```{r, warning=FALSE}
# Assuming your dataframe is called drinks_data_combined
drinks_data_combined <- mutate_all(drinks_data_combined, ~ifelse(. == "NA", NA, .))
# Replace "your_column_names" with the actual names of columns 7 to 310 in your dataframe
columns_to_modify <- colnames(drinks_data_combined)[7:310]

# Use mutate_at to apply the string replacement to the selected columns
drinks_data_combined <- mutate_at(drinks_data_combined, vars(columns_to_modify), funs(gsub("\\bbacardi\\b", "", .)))
#removing whitespace in column names
colnames(drinks_data_combined) <- colnames(drinks_data_combined) %>% str_trim()
```

The binary columns have been pasted at the end of the dataframe. A good identifier to separate them is that they contain the word ingredient in the name (you may use grepl).

```{r}
#dropping one NA in type and one NA in recipe
cocktails <- drinks_data_combined %>% 
  filter(!is.na(Type) & ! is.na(Recipe))
```

#Checking alcoholic drinks
```{r}
all_cols <- unique(colnames(cocktails)) %>% str_trim() %>% sort()

categorical_vars <- c("Name", "Type", "Category", "Picture", "Glass", "Recipe")

unique_ing_vec <- setdiff(all_cols, categorical_vars)

alcohol_vec <- c("151 bacardi rum", "absinthe", "absolut citron", "absolut kurant", "absolut peppar", "absolut vodka", "advocaat", "ale", "amaretto", "anisette", "aperol", "apfelkorn", "apple brandy", "apple cider", "apple schnapps", "applejack", "apricot brandy", "aquavit", "añejo rum", "bacardi limon", "baileys irish cream", "banana liqueur", "beer", "benedictine", "black sambuca", "blackberry brandy", "blended whiskey", "blue curacao", "blueberry schnapps", "bourbon", "brandy", "butterscotch schnapps", "cachaca", "campari", "chambord raspberry liqueur", "champagne", "cherry brandy", "cherry heering", "cherry liqueur","chocolate liqueur", "cider", "coconut liqueur", "coconut rum", "coffee brandy", "coffee liqueur", "cognac", "cointreau", "corona", "cranberry vodka", "cream of coconut", "creme de banane", "creme de cacao", "creme de cassis", "creme de mure", "crown royal", "curacao", "dark creme de cacao", "dark rum","drambuie", "dry vermouth", "dubonnet rouge", "erin cream", "everclear", "firewater", "frangelico", "galliano", "gin", "godiva liqueur", "gold tequila", "goldschlager", "grain alcohol", "grand marnier", "green chartreuse", "green creme de menthe", "guinness stout", "hot damn", "irish cream", "irish whiskey", "jack daniels", "jim beam", "johnnie walker", "jägermeister", "kahlua", "kirschwasser", "kiwi liqueur", "lager", "lemon vodka", "light rum", "lillet blanc", "lime vodka", "malibu rum", "maraschino liqueur", "melon liqueur", "midori melon liqueur", "orange curacao", "ouzo", "peach brandy", "peach schnapps", "peach vodka", "peppermint schnapps", "peychaud bitters", "pisang ambon", "pisco", "port", "prosecco", "raspberry liqueur", "raspberry vodka", "red wine", "ricard", "rum", "rumple minze", "rye whiskey", "sambuca", "scotch", "sherry", "sloe gin", "southern comfort", "spiced rum", "st. germain", "strawberry liqueur", "strawberry schnapps", "sweet vermouth", "tennessee whiskey", "tequila", "tia maria", "triple sec", "vanilla vodka", "vermouth", "vodka", "whiskey", "white creme de menthe", "white rum", "wild turkey", "wine", "yellow chartreuse", "yukon jack")

side_ing_vec <- setdiff(unique_ing_vec, alcohol_vec)
```

We have 8 cocktails that are marked as alcoholic but don't contain alcohol in the vector
```{r}
filtered_cocktails <- cocktails %>%
  mutate(sums = rowSums(!is.na(.[, alcohol_vec]))) %>% 
  select(sums, everything()) %>% 
  filter(sums !=0)
#467 of type alcoholic and 8 of optional alcohol
filtered_cocktails %>% 
  count(Type)
#--> 475 which contain alcohol
contain_alc <- filtered_cocktails %>% pull(Name)

#486 which are of either type --> 11 which are missing
are_alc <- cocktails %>% 
  filter(Type =="alcoholic" | Type == "optional alcohol")

# out of those 11, we had only 3 names which contained at least one ingredient
are_alc %>%
  filter(!Name %in% contain_alc) %>% 
  select(-c(Category, Picture, Glass, Recipe)) %>%
  pivot_longer(-c(Name, Type), names_to = "ingredient", values_to = "quantity", values_drop_na  = T)

alcohol_vec <- append(alcohol_vec, c("maui","peachtree schnapps","surge")) %>% sort()
```
Dropping cocktails which contain no ingredients, or those that don't make sense:
- black forest shake --> alcoholic but contains only ice
- egg-nog - classic cooked --> optional alcohol but without alcohol ingredient listed
- zoksel - alcoholic type but contains only coca-cola
```{r}
#filtering out the 3 cocktails in the list above
cocktails <- cocktails %>% 
  filter(!Name %in% c("black forest shake", "egg-nog - classic cooked", "zoksel"))

#removing additional 7 cocktails which contain 0 ingredients
cocktails <- cocktails %>% 
  mutate(sums = rowSums(!is.na(.[, unique_ing_vec]))) %>%
  filter(sums != 0) %>% 
  select(-sums)
```

# Conversion function
In this code we will create a function for transforming the units of the ingredients from imperial to international system. This will be incorporated as a function for the R package afterwards. The methodology for making the conversion is the following: We will upload the data, look at the information that contains the measurements. If the unit contains imperial units (for example ounces - oz) we will convert them to a practical unit in the international system (for example ml).

We start by separating the string that contains the information for the measurements between number and unit. We intend to see how many different units we would need to pass from imperial to international.

```{r}
# Assuming your dataframe is named df
library(stringr)

# Specify the range of columns
measure_cols <- cocktails[, 7:310]

# Extract unique measurement units from all cells in the specified range
unique_units <- unique(unlist(apply(measure_cols, c(1, 2), function(x) {
  units <- str_extract(x, "[a-zA-Z]+")
  units[!is.na(units)]  # Remove NAs
})))

```

```{r}
# Print the unique measurement units
print(unique_units)
```

```{r}
imperial<- c("oz", "gal", "qt","pint", "lb", "quart","inch")
international <- c("cl","ml","dl","l","gr","kg")
```

The logical step would be making the conversion only for those units in imperial or international. While analyzing the data we realized that the numbers are not prepared for working in a conversion of units. the main issue is the following: we have high heterogeneity in the way in which numbers are presented. The following are the main options that have been spotted as examples: (1) 1 oz; (2) 1.5 lb; (3) 1/4 kg; (4) 1 1/2 spoon; (5) 1-2 dash. We need to create one single format to show the numbers, preferably as decimal or whole numbers, and not in fractional notation or as a range. For this purpose we implement the following function for homogenizing the numbers:

```{r}
correctFractions <- function(data, decimals = 1) {
  # Iterate over the rows and columns of the dataset
  for (i in 1:nrow(data)) {
    for (j in 1:ncol(data)) {
      # Check if the value follows the specified pattern for integers plus fraction
      integerPattern <- "([0-9]+) ([0-9]+)/([0-9]+)"
      if (!is.na(data[i, j]) && grepl(integerPattern, data[i, j])) {
        # Extract parts of the value
        parts <- str_match(data[i, j], integerPattern)
        integerPart <- as.numeric(parts[2])
        numerator <- as.numeric(parts[3])
        denominator <- as.numeric(parts[4])
        
        # Calculate the sum
        sumValue <- integerPart + numerator/denominator
        
        # Round the value to the specified number of decimals
        roundedSum <- round(sumValue, decimals)
        
        # Create the new value without repeating the unit
        new_value <- paste(roundedSum, gsub("\\d+ \\d+/\\d+ (.+)", "\\1", data[i, j]))
        
        # Update the dataset
        data[i, j] <- new_value
      } else {
        # Check if the value follows the specified pattern for fractions
        fractionPattern <- "([0-9]+)/([0-9]+)"
        if (!is.na(data[i, j]) && grepl(fractionPattern, data[i, j])) {
          # Extract parts of the value
          parts <- str_match(data[i, j], fractionPattern)
          numerator <- as.numeric(parts[2])
          denominator <- as.numeric(parts[3])
          
          # Calculate the decimal value
          decimalValue <- numerator/denominator
          
          # Round the value to the specified number of decimals
          roundedValue <- round(decimalValue, decimals)
          
          # Create the new value
          new_value <- paste(roundedValue, gsub("\\d+/\\d+ (.+)", "\\1", data[i, j]))
          
          # Update the dataset
          data[i, j] <- new_value
        } else {
          # Check if the value follows the specified pattern for "number-dash-number"
          dashPattern <- "([0-9]+)-([0-9]+)"
          if (!is.na(data[i, j]) && grepl(dashPattern, data[i, j])) {
            # Extract the part after the dash
            parts <- str_match(data[i, j], dashPattern)
            afterDash <- parts[3]
            
            # Create the new value
            new_value <- paste(afterDash, gsub("\\d+-\\d+ (.+)", "\\1", data[i, j]))
            
            # Update the dataset
            data[i, j] <- new_value
          }
        }
      }
    }
  }
  
  # Remove any values of the type "space-number-slash-number"
  data <- data.frame(lapply(data, function(x) gsub("\\s\\d+/\\d+", "", x, perl = TRUE)))
  
  # Clean column names and remove trailing whitespaces
  colnames(data) <- gsub("\\.$", "", gsub("\\.", " ", colnames(data)))
  colnames(data) <- gsub("\\s+$", "", colnames(data))
  
  # Print the final version of the dataset
  return(data)
}

# Example usage:
# df <- correctFractions(df, decimals = 1)


# Function to clean column names (in case the inside function does not work in R package)
# clean_column_names <- function(df) {
#   colnames(df) <- gsub("\\.$", "", gsub("\\.", " ", colnames(df)))
#   return(df)
# }

# Apply the function to the data frame
cocktails <- correctFractions(cocktails, decimals = 2)

# Print the cleaned data frame
print(cocktails)
```

#Creating rda files
```{r eval=FALSE}
usethis::use_data(cocktails, overwrite = TRUE)
usethis::use_data(unique_ing_vec, alcohol_vec, side_ing_vec, internal = TRUE, overwrite = TRUE)
# usethis::use_data(unique_ing_vec)
# usethis::use_data(alcohol_vec, overwrite = TRUE)
# usethis::use_data(side_ing_vec)
```

