---
title: "R packages functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{R packages functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

While the Cocktail Selector Shiny app offers an interactive experience, the underlying functionality is powered by R functions that can be leveraged directly. This article details how you can utilize these functions independently of the Shiny app interface.
Below, we detail the various functions provided by the CocktailSelector package along with descriptions of their individual uses.

# Filerting function 

This function is used to create the list of drinks based on inputs from the dropdowns.
It filters the datasets based on one alcohol and two side ingredients and returns a filtered dataframe with drinks corresponding to those criteria.

*Parameters :* 
- `df`: The cocktails dataframe included in the package

- `alcohol`: A string which is in the alcohol_Vec vector

- `ing1`: A string which is in the side_ing_vec vector

- `ing2`: A string which is in the side_ing_vec vector

For example : 

```{r, eval=FALSE}
library(CocktailSelector)
cocktail_with_vodka_and_cranberry <- filterCocktails(cocktails, "vodka", "cranberry juice")
print(cocktail_with_vodka_and_cranberry)
```
It will return a filtered drinks dataframe based on inputs(vodka and cranberry juice)


# Ingredients table function

This function takes as input the cocktails dataframe and the name of a cocktail in that dataframe and returns its ingredients along with quantities.

*Parameters :* 

- `df`: The cocktails dataframe from the package

- `selected_cocktail`: A cocktail name string

```{r, eval=FALSE}
library(CocktailSelector)
daiquiri <- ingredientsTable(cocktails, "daiquiri")
print(daiquiri)
```
It will return an HTML table with ingredients and corresponding quantities


# The suprise me function
The function can be used to randomly select a cocktail from the dataframe. The output can the be used in the {ingredientsTable} function in order to get the ingredients and quantities needed for the cocktail.

*example :* 

```{r, eval=FALSE}
# Using it on its own
surpriseMe()

#using it with ingredientsTable
ingredientsTable(selected_cocktail = surpriseMe()$Name)
```

It will return a randomly selected cocktail within the {cocktails} dataframe
